---
title: "2021 DSO 545 Spring Midterm Exam"
date: "3/11/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#Install the following packages:
#ggplot2
#gcookbook
#dplyr
#dslabs
#gridExtra
#lme4
```{r}
library(ggplot2)
library(gcookbook)
library(dplyr)
library(dslabs)
library(gridExtra)
library(lme4)
```


1. (5 points) Within Rstudio there is a dataset called cabbage_exp, save it as an object called cabbagedf. Then create the visualization below with the Cultivar and Weight variables and fill by Cultivar. Generate and replicate the graph below.
```{r}
cabbagedf = cabbage_exp

cabbagedf %>% 
  group_by(Cultivar) %>%
  summarise(Weight = sum(Weight)) %>%
  ggplot(aes(x=Cultivar, y=Weight, fill=Cultivar)) + 
  geom_col()
```


2. (5 points) Now generate the multiple line graph as shown below with the  cabbagedf. Generate and replicate the graph below.
```{r}
cabbagedf %>%
  ggplot(aes(x=Date, y=Weight, group=Cultivar, colour=Cultivar)) + 
  geom_line()
```


3. (5 points) Now create a new dataframe called ce that has cabbbage_exp using dplyr verb and with a new variable called percent_weight that entails the calculation of the percent weight grouped by date. Then check to see that the new variable is in the dataframe.
```{r}
ce = cabbagedf %>%
  group_by(Date) %>%
  mutate(percent_weight = Weight / sum(Weight))
ce

```


4. (5 points) After computing the new column, make a stacked bar chart with the ce dataframe and the percent_weight variable and fill by Cultivar. Generate and replicate the graph below.
```{r}
ce %>%
  ggplot(aes(x=Date, y=percent_weight, fill=Cultivar)) +
  geom_col()
```


5. (5 points) Make a stacked bar graph that shows proportions of weight per date. Generate and replicate the graph below. 
```{r}
ce %>%
  ggplot(aes(x=Date, y=Weight, fill=Cultivar)) +
  geom_col()
```
 
 
6.(5 points) New dataset (within R), now obtain the us_contagious_diseases data by typing in us_contagious_diseases and save it in an object called uscdf. Generate and replicate the graph shown below that depicts the range of each of contagious disease throughout different periods. In order to make it look like the graph shown, include the following line of code in your ggplot: scale_fill_discrete(name = “Disease”).
```{r}
uscdf = us_contagious_diseases

uscdf %>%
  ggplot(aes(y=disease, x=year, fill=disease)) + 
  geom_boxplot() + 
  scale_fill_discrete(name = "Disease") + 
  ggtitle("Disease Over the Years") + 
  ylab("Disease") + 
  xlab("Year")
```


7. (5 points) Obtain the information for “Measles” that had a count greater than 0 for the year “2000” and make sure to not include any missing values from the population variable.Then create a variable called rate which will be the result of count/population multiplied by 10000 and multiplied by 52 divided by the weeks_reporting. Create the plot that shows the rate per state and reorder it so that the it shows the state with the highest rates at the top of the graph. Generate and replicate the graph shown below.
```{r}
uscdf %>%
  filter(disease == "Measles", count > 0, year == 2000, !is.na(population)) %>%
  mutate(rate = count / population * 10000 * 52 / weeks_reporting) %>%
  ggplot(aes(x=rate, y=reorder(state, rate))) + 
  geom_col()
```


8.(5 points) Copy the code from the previous questions and  enhance your plot. Make sure to add the appropriate x axis label and change the y label to “State” and a title “Measles Rate by State” and finally change the color to a color of the graph to a color of your choice. Obtain the information for “Measles” for the year “2000” and make sure to not include any missing values. Generate and replicate the graph shown below.
```{r}
uscdf %>%
  filter(disease == "Measles", count > 0, year == 2000, !is.na(population)) %>%
  mutate(rate = count / population * 10000 * 52 / weeks_reporting) %>%
  ggplot(aes(x=rate, y=reorder(state, rate))) + 
  geom_col(fill="blue") + 
  ggtitle("Measles Rate by State") + 
  ylab("State") + 
  xlab("Rate")
```


9.(5 points) New dataset, within R use the iris dataset. Check to see if the correlations between each combination of variables within the iris dataset excluding the categorical variable Species. Save the iris dataset without the Species variable in an object called numiris (show your code). Run all the correlations.
```{r}
numiris = iris %>% select(-Species)

cor(numiris)
```


10. (5 points) Which pairing is the has the strongest positive correlation? Then use ggplot to visualize it with the appropriate geom type of graph to display the strongest positive correlation.
```{r}
# Petal.Length and Petal.Width
numiris %>%
  ggplot(aes(x=Petal.Length, y=Petal.Width)) + 
  geom_point() + 
  geom_smooth(method=lm)
```


11.(5 points) Within R there is a dataset called sleepstudy, go ahead and save it as an object in your global environment as sleepstudydf. Then generate and replicate  the visualization below.
```{r}
sleepstudydf = sleepstudy

sleepstudydf %>%
  ggplot(aes(x=Days, y=Reaction, group=Days)) + 
  geom_boxplot()
```


12.(5 points) Within R there is the quakes dataset, save it as an object called earthquakedf. Convert the magnitude into the following categories: anything below a 5 = light, anything between 5 and 5.9 = moderate, and anything 6 or above = strong. Save categorized mag as magcoded. Check the class and make sure it is a factor variable.
```{r}
earthquakedf = quakes

earthquakedf = mutate(earthquakedf, magcoded = factor(ifelse(mag < 5, "light", ifelse(between(mag, 5, 5.9), "moderate", "major"))))

class(earthquakedf$magcoded)
```


13.(5 points) Then visualize the newly categorized variable of magcoded with stations variable, choose the most appropriate graph. Then reorder the visualization so as to generate and replicate the graph below.  (to replicate the graph below without the legend, incorporate this line of code at the end of your ggplot: theme(legend.position = “none”))
```{r}
earthquakedf %>%
  ggplot(aes(x=magcoded, y=stations)) + 
  geom_col() + 
  theme(legend.position = "none")

earthquakedf %>%
  ggplot(aes(x=magcoded, y=stations, fill=magcoded)) + 
  geom_col() + 
  theme(legend.position = "none") + 
  ylab("Number of Stations") + 
  xlab("Magnitude Intensity")
```


14.	(5 points) Now save the gapminder dataset as an object called gadp. Then create a subset that includes only the following variables: country, year, fertility, population, gdp, continent, region that have a gdp greater than or equal to 10 billion (10000000000).
```{r}
gadp = gapminder

gadp2 = gadp %>% 
  select(country, year, fertility, population, gdp, continent, region) %>% 
  filter(gdp >= 10^10)
```


15.	(5 points) Now create a plot that examines the distribution of fertility by using a histogram overlayed with a density curve. Generate and replicate the graph below using two colors of your choice to represent the histogram and the density curve.
```{r}
gadp2 %>%
  ggplot(aes(x=fertility)) +
  geom_histogram(aes(y=..density..), fill="#00BFC4") + 
  geom_density(alpha=.6, fill="#F8766D")
```


16.	(5 points) Create a boxplot of the population by continent. Then color by continent to distinguish between continents. Type options(scipen = 999) prior to running your ggplot code so the Population values don’t show up in scientific notation.
```{r}
options(scipen = 999)
gadp2 %>%
  ggplot(aes(x=continent, y=population, color=continent)) + 
  geom_boxplot() + 
  ggtitle("POPULATION BY CONTINENTS") +
  ylab("POPULATION") + 
  xlab("CONTINENTS")
```


17.	(5 points) Using the piping operator and appropriate verb choose the following variables: country and population. Then specify the country to obtain the mean of “Thailand” and the mean of “Germany” and and save it an object called, Average_pop that should have two means.
```{r}
Average_pop = gadp2 %>%
  select(country, population) %>%
  filter(country %in% c("Thailand", "Germany")) %>%
  group_by(country) %>%
  summarise(m = mean(population)) %>%
  pull(m)
Average_pop
```


18.	(5 points) Obtain the summary of the gadp population variable and create labels of ‘lowpop’, ‘modpop’, and ‘highpop’ based on anything less than or equal to the 1st quartile as lowpop, anything from lowpop up and including the 3rd quartile as ‘modpop’ and anything above as ‘highpop’ (use the summary function to find the quartiles) and save it in an object called popfactor. In order to see the value and remove the scientific notation use the function options(scipen = 999).
```{r}
options(scipen = 999)

summary(gadp2$population)  # 1st Qu.: 7399075; 3rd Qu.: 51128780

popfactor = gadp2 %>% 
  mutate(popfactor = ifelse(population <= 7399075, "lowpop", ifelse(population > 51128780, "highpop", "modpop"))) %>%
  pull(popfactor)
```


19. (5 points) Convert the popfactor object to an actual factor variable and have it be an extra variable in the gadp dataset, and finally graph it by country.
```{r}
popfactor = factor(popfactor)

gadp2$popfactor = popfactor

# Not quite understand "finally graph it by country"
gadp2 %>%
  ggplot(aes(x=popfactor)) +
  geom_bar()
```


20. (5 points) Generate and replicate the graph below.
```{r}
tmp = data.frame(
  continent=c("Africa", "Africa", "Africa", "Americas", "Americas", "Americas", "Asia", "Asia", "Asia", "Europe", "Europe", "Europe", "Oceania", "Oceania"), 
  popfactor=c("Low", "Mod", "High", "Low", "Mod", "High", "Low", "Mod", "High", "Low", "Mod", "High", "Mod", "High"), 
  value=c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 5, 5))

tmp$popfactor = factor(tmp$popfactor, c("Low", "Mod", "High"))

tmp %>%
  ggplot(aes(x=popfactor, y=value, fill=continent)) +
  geom_col(position='dodge') + 
  ggtitle("Levels of Population by Continents") +
  ylab("Continent") + 
  xlab("Population") + 
  theme(axis.ticks.y=element_blank(), 
        axis.text.y=element_blank(),
        panel.grid.minor.y=element_blank(),
        panel.grid.major.y=element_blank())
```

